\textcolor{blue}{A gradient-based optimization method for motion planning}$^{1}$

### 2016-01-14 paper.tex
Fix Figures typo, add algorithm description, update RS and PRS

In this section, the GB path-optimizer Algorithm~\ref{algo:gradient} is described
according to the previous steps definitions.
\textcolor{blue}{PARTIR A FINIR!!}
\textcolor{blue}{Remind that, the QP optimal step $\p$ computed by \texttt{computeIterate} is known.
The idea of the algorithm is to process ....
So the difficulty here is to handle 
the scalar parameter $\alpha$ determining how much of the computed step  will be traveled along. 
As presented in Algorithm~\ref{algo:gradient}, $\alpha$ takes two values, $\alpha_{init}<1$ to process smaller steps or $1$ to go directly to the optimum. This latter case is interesting since, if this optimal path is collision-free, the algorithm has converged and returns the path as the solution.
Choosing to travel small steps from a valid path decrease the chances of being in collision. Besides if one occurs, the collision-constraint is computed on the last valid path.
$\alpha_{init}$ could be computed from geometrical considerations.
...
smaller steps toward the minimum are iteratively applied and tested, and constraints are added each time a collision is found with a reduced step (i.e. $\alpha \neq 1$).}

The collision detection on a path is handled by \texttt{validatePath}, which returns $true$ if the given path is collision-free.

@@

This part gathers optimization results performed on the planning software 
Humanoid Path Planner~\cite{hpp}. Initial paths are obtained with two 
kinds of probabilistic planners: Visibility-PRM~\cite{visibility-prm} and 
RRT-connect~\cite{rrt-connect}. Unless another value is provided, $\alpha_{init}$ 
is set to 0.2.

@@

\begin{algorithm}
\begin{algorithmic}%[1] % line numerotation
\INPUT{Path to optimize $\xx$}
\OUTPUT{Optimized collision-free path $\xx$}
   \State $nbFailures$ $\gets 0$
   \While {$nbFailures < maxNbFailures$}
        \State $failure \gets true$
        \State $t_1 < t_2 \gets \mbox{random numbers in }[0,1]$
        \State $lp0 \gets \texttt{steeringMethod} (\xx(0), \xx(t_1))$
        \State $lp1 \gets \texttt{steeringMethod} (\xx(t_1), \xx(t_2))$
        \State $lp2 \gets \texttt{steeringMethod} (\xx(t_2), \xx(1))$
        \State $newPath \gets \mbox{empty path defined on }[0,0]$
   	\If{\texttt{validatePath($lp0$)}}
          \State $newPath \gets lp0$; $failure \gets false$
        \Else
          \State $newPath \gets \xx_{|[0,t_1]}$
        \EndIf
   	\If{\texttt{validatePath($lp1$)}}
          \State $newPath \gets \texttt{concatenate} (newPath, lp1)$
          \State $failure \gets false$
        \Else
          \State $newPath \gets \texttt{concatenate} (newPath, \xx_{|[t_1,t_2]})$
        \EndIf
   	\If{\texttt{validatePath($lp2$)}}
          \State $newPath \gets \texttt{concatenate} (newPath, lp2)$
          \State $failure \gets false$
        \Else
          \State $newPath \gets \texttt{concatenate} (newPath, \xx_{|[t_2,1]})$
        \EndIf
        \State $\xx \gets newPath$
        \If {$failure$} $nbFailures \gets nbFailures + 1$
        \EndIf
      \EndWhile
    \Return $\xx$
\end{algorithmic}
\caption{Random shortcut as adapted from~\cite{Sekhavat-Svestka1998} 
Section~6.4.1. \texttt{steeringMethod} returns the linear interpolation between 
two configurations. $\xx_{|I}$ denotes path $\xx$ restricted to interval $I$. 
$maxNbFailures$ is a parameter that affects time of computation and quality of 
the result.}\label{algo:random-shortcut}
\end{algorithm}

@@

We also experiment our algorithm on more complex robots and environments.
In the included video and Figures~\ref{pr2_final},~\ref{fiad} 
and~\ref{fig:trajectories}, we present multiple situations where the GB algorithm 
is tested and compared to RS~\cite{randomShortcutHPP}
Algorithm~\ref{algo:random-shortcut} and PRS~\cite{partialrandomShortcutHPP} 
adapted from~\cite{GeraertsIJRR07}.
The termination condition of RS allows it to try 
shortening the path until five iterations of non-improvement are reached. 
This corresponds to $maxNbFailures$ in Algorithm~\ref{algo:random-shortcut}.
For our implementation of PRS, first a shortcut is tried on each DOF between 
$\conf_0$ and $\conf_{wp+1}$. 
Then, for each DOF, configurations are sampled as in 
Algorithm~\ref{algo:random-shortcut}. The steering-method returns a path made of 
an interpolation on only 
the current DOF and based on the previous subpath for other DOF. If this path is 
collision-free, we replace it as in RS. This process is also stopped when five 
iterations of non-improvement are reached.

>>>>

In this section, the GB path-optimizer Algorithm~\ref{algo:gradient} is described
according to the previous steps definitions.
\textcolor{blue}{Remind that, the QP optimal step $\p$ computed by \texttt{computeIterate} is known.
The idea of the algorithm is to process iterations that reduce the path length according to the QP cost. If collision-constraints have been added to the QP, further iterations will comply with them. The algorithm stops whether the QP minimum is reached and collision-free. The collision detection on a path is handled by \texttt{validatePath}, which returns $true$ if the given path is collision-free.
}

\textcolor{blue}{
The main difficulty is to handle 
the scalar parameter $\alpha$ determining how much of the computed step  will be traveled along. 
As presented in Algorithm~\ref{algo:gradient}, $\alpha$ takes two values, $\alpha_{init}<1$ to process smaller steps or $1$ to go directly to the optimum. This latter case is interesting since, if this optimal path is collision-free, the algorithm has converged and returns the path as the solution.
Choosing to travel small steps from a valid path decrease the chances of being in collision. Besides if one occurs, the collision-constraint is computed on the last valid path, which is not too much deformed compared to the path that has collisions. As a result, collision-constraints are only computed (\texttt{computeCollisionConstraint}) and added (\texttt{addCollisionConstraint}) when performing a reduced iteration (i.e. $\alpha = \alpha_{init}$).
}

@@

This part gathers optimization results performed on the planning software 
Humanoid Path Planner~\cite{hpp}. Initial paths are obtained with two 
kinds of probabilistic planners: Visibility-PRM~\cite{visibility-prm} and 
RRT-connect~\cite{rrt-connect}. Unless another value is provided, $\alpha_{init}$ 
is set to 0.2. \textcolor{blue}{A further section provides a discussion on the $\alpha_{init}$ value setting.}

@@

\subsubsection{Analysis of $\alpha_{init}$ influence}
\textcolor{blue}{Results without comparison to RS and PRS here
$\alpha_{init}$ could be computed from geometrical considerations.
}
TODO !!

\subsubsection{Comparison to random shortcut algorithms}
\begin{algorithm}
\begin{algorithmic}%[1] % line numerotation
\INPUT{Path to optimize $\xx$, time limit $t_{lim}$}
\OUTPUT{Optimized collision-free path $\xx$}
   \State \textcolor{blue}{$t_{start} \gets \texttt{currentTime()}; \ t \gets 0$}
   \While {\textcolor{blue}{$t < t_{lim}$}}
        \State $failure \gets true$
        \State $t_1 < t_2 \gets \mbox{random numbers in }[0,1]$
        \State $lp0 \gets \texttt{steeringMethod} (\xx(0), \xx(t_1))$
        \State $lp1 \gets \texttt{steeringMethod} (\xx(t_1), \xx(t_2))$
        \State $lp2 \gets \texttt{steeringMethod} (\xx(t_2), \xx(1))$
        \State $newPath \gets \mbox{empty path defined on }[0,0]$
   	\If{\texttt{validatePath($lp0$)}}
          \State $newPath \gets lp0$;
        \Else
          \State $newPath \gets \xx_{|[0,t_1]}$
        \EndIf
   	\If{\texttt{validatePath($lp1$)}}
          \State $newPath \gets \texttt{concatenate} (newPath, lp1)$
        \Else
          \State $newPath \gets \texttt{concatenate} (newPath, \xx_{|[t_1,t_2]})$
        \EndIf
   	\If{\texttt{validatePath($lp2$)}}
          \State $newPath \gets \texttt{concatenate} (newPath, lp2)$
        \Else
          \State $newPath \gets \texttt{concatenate} (newPath, \xx_{|[t_2,1]})$
        \EndIf
        \State $\xx \gets newPath$
        \State \textcolor{blue}{$t \gets \texttt{currentTime()} - t_{start}$}
      \EndWhile
    \Return $\xx$
\end{algorithmic}
\caption{Random shortcut as adapted from~\cite{Sekhavat-Svestka1998} 
Section~6.4.1. \texttt{steeringMethod} returns the linear interpolation between 
two configurations. $\xx_{|I}$ denotes path $\xx$ restricted to interval $I$. 
\textcolor{blue}{$t_{lim}$ represents the duration limitation of the algorithm.}}\label{algo:random-shortcut}
\end{algorithm}

@@

We also experiment our algorithm on more complex robots and environments.
In the included video and Figures~\ref{pr2_final},~\ref{fiad} 
and~\ref{fig:trajectories}, we present multiple situations where the GB algorithm 
is tested and compared to RS~\cite{randomShortcutHPP}
Algorithm~\ref{algo:random-shortcut} and PRS~\cite{partialrandomShortcutHPP} 
adapted from~\cite{GeraertsIJRR07}.
\textcolor{blue}{
The termination condition of RS allows it to try 
shortening the path until the optimization duration overflows a time limit. 
This corresponds to $t_{lim}$ in Algorithm~\ref{algo:random-shortcut} and is typically equal to the Gradient-Based optimization duration.
%For our implementation of PRS, first a shortcut is tried on each DOF between $\conf_0$ and $\conf_{wp+1}$. % NOT ANYMORE
The implementation of PRS is identical to~\cite{GeraertsIJRR07}:
for a random DOF, configurations are sampled as in 
Algorithm~\ref{algo:random-shortcut}. The steering-method returns a path made of 
an interpolation on only 
the current DOF and based on the previous subpath for other DOF. If this path is 
collision-free, we replace it as in RS. As RS, the process is stopped when the duration exceeds $t_{lim}$.
}
